---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Read activity.csv hat had been extract at local directory
file is downloaded in url https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip

```{r echo = TRUE}
read_df <- read.csv("activity.csv", header = T, sep = ",")
```

set date format `"%Y-%m-%d"`

```{r echo = TRUE}
read_df$date <- as.Date(read_df$date, format = "%Y-%m-%d")
```
## What is mean total number of steps taken per day?

**Agregate** by total steps and day with naming column **'Steps' and 'Date'**

```{r echo = TRUE}
#act.day <- aggregate(read_df$steps, by = list(read_df$date), sum )
act.day <- aggregate(steps ~ date, read_df, sum)
head(act.day)
names(act.day)[2] <- "Steps"
names(act.day)[1] <- "Date"
head(act.day)

```
**Histogram total number of step per day**

```{r echo = TRUE}
hist(act.day$Steps, col = 2, main = "Histogram number of steps taken each day", xlab = "Total number of steps taken each day")
```

**Calculate mean and median of the total number of steps taken per day**
```{r echo = TRUE}
mean(act.day$Steps)
median(act.day$Steps)

```

## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r echo = TRUE}
act.interval <- aggregate(steps ~ interval, read_df, mean)
head(act.interval)
names(act.interval)[2] <- "Steps"
names(act.interval)[1] <- "Interval"
head(act.interval)

plot(act.interval$Interval, act.interval$Steps, type='l', col=2, 
     main="Average number of steps mean over all days", xlab="Interval", 
     ylab="Average number of steps")
```

Find which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

```{r echo = TRUE}
#find row number that has maximum average number of steps in an interval
row_number <- which.max(act.interval$Steps)
row_number

#Get data from row number base on 'row_number'
table_interval_steps <- act.interval[row_number,]
table_interval_steps
```

## Imputing missing values

```{r echo = TRUE}
#calculate NA 
read_NA <- sum(is.na(read_df$steps))
read_NA

# perform the imputation
for (i in 1:nrow(read_df)){
  if (is.na(read_df$steps[i])){
    interval_val <- read_df$interval[i]
    row_id <- which(act.interval$Interval == interval_val)
    steps_val <- act.interval$Steps[row_id]
    read_df$steps[i] <- steps_val
  }
}

# Agregate by total steps and day with  NA imputation 'new dataset "act.day.imputed"' 
act.day.imputed <- aggregate(steps ~ date, read_df, sum)
head(act.day.imputed)

# Histogram of imputed NA total number of steps per day 
hist(act.day.imputed$steps, col = 2, main = "(Imputed) Histogram number of steps taken each day", xlab = "Total number of steps taken each day")

# get mean and media of total number of steps per day with imputed NA
mean(act.day.imputed$steps)
median(act.day.imputed$steps)

# get mean and media of total number of steps per day without imputed NA
mean(act.day$Steps)
median(act.day$Steps)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r echo = TRUE}
# Add new column that define weekday or weekend to dataset
read_df$day <- weekdays(read_df$date)

# Add initilize to column
read_df$day_type <- c("weekday")

# Define Saturday and sunday as weekend
for (i in 1:nrow(read_df)){
  if (read_df$day[i] == "Sabtu" || read_df$day == "Ahad"){
    read_df$day_type[i] <- "Weekend"
  }
}

# convert day_time from character to factor
read_df$day_type <- as.factor(read_df$day_type)

# aggregate steps as interval to get average number of steps in an interval across all days
table_interval_steps_imputed <- aggregate(steps ~ interval+day_type, read_df, mean)

library(ggplot2)

qplot(interval, steps, data=table_interval_steps_imputed, geom=c("line"), xlab="Interval", ylab="Number of steps", main="") + facet_wrap(~ day_type, ncol=2)


```

